\chapter{Broader Impacts}\label{chap:broader-impacts}

ChromaGuide is intended as a \emph{computational decision-support} framework for prioritizing \gRNA{}s in CRISPR-Cas9 workflows. Its broader impact derives from integrating chromatin context into prediction and from reporting calibrated uncertainty, which together can improve downstream decision making under biological shift.

\section{Translational relevance and decision support}

By combining sequence with epigenomic accessibility/state features, ChromaGuide aims to reduce failed experiments caused by locus-specific chromatin effects and to improve robustness when the deployment cell context differs from the training data. Outputs are intended for \emph{guide triage and hypothesis generation} (not clinical decision making) and should be used alongside standard validation, safety review, and domain expertise.

\section{Scientific and methodological impact (reproducibility and evaluation)}

A central contribution is a leakage-controlled evaluation protocol (gene-held-out plus dataset/cell-line-held-out stress tests) that makes failure modes under biological shift explicit. Conformal prediction intervals add an outcome that can be audited quantitatively (empirical coverage and interval width), complementing standard ranking metrics.

\section{Responsible use, governance, and limitations}

\begin{itemize}
\item \textbf{Reliability and communication:} Report uncertainty (coverage, \ECE{}, interval width) and clearly state the split/protocol; avoid overclaiming accuracy outside evaluated domains.
\item \textbf{Representativeness and bias:} Public datasets over-represent common cell lines and protocols; therefore, report performance under cross-domain stress tests and avoid implying general clinical validity without evidence.
\item \textbf{Data governance:} For patient-derived or proprietary screens, document provenance, permitted uses, and what is excluded from release.
\item \textbf{Dual-use awareness:} Any release should include a scope statement, limitations, and guidance against harmful use; controlled distribution of trained models is an option if risk warrants.
\item \textbf{Non-causality and shift limits:} Learned associations are not causal, and strong batch effects/unmeasured confounders can degrade calibration even with weighted conformal methods; these are treated as first-class failure modes to monitor.
\end{itemize}

\section{Conclusion}

Overall, ChromaGuide aims to make \gRNA{} prioritization more context-aware and more transparent about uncertainty, while emphasizing evaluation protocols that support reproducible, shift-aware reporting.
